---
title: "Reproducible Research Project 1"
author: "Sanya Sharma"
date: "July 21, 2018"
output: html_document
---

## Reproducible Research project 1


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Reading Data

```{r simulation, echo=FALSE}

activity <- read.csv("activity.csv")
dates <- strptime(activity$date, "%Y-%m-%d")
activity$date <- dates
Dates <- unique(dates)
Intervals <- unique(activity$interval)
```
## Part1- Histogram of total steps taken in a day

##Histogram
```{r}
steps <- split(activity$steps, dates$yday)
totalStepsPerDay <- sapply(steps, sum, na.rm=TRUE)
plot(Dates, totalStepsPerDay, main="Histogram of steps taken each day", 
     xlab="Date (October to November 2012)", ylab="Frequency", type="h", lwd=4, col="red")

```
## Part2- Mean and Median of steps per day

##The mean steps per day 
```{r}
StepsPerDaymean <- sapply(steps, mean, na.rm=TRUE)
mean <- data.frame(date=Dates, StepsPerDaymean=StepsPerDaymean, row.names=NULL)
mean
```

##The median steps per day 

```{r}
StepsPerDaymedian <- sapply(steps, median, na.rm=TRUE)
median <- data.frame(date=Dates, StepsPerDaymedian=StepsPerDaymedian, row.names=NULL)
median
```
## Part3- Split up the data according to the interval

```{r}

interval <- split(activity$steps, activity$interval)

StepsPerIntervalaverage <- sapply(interval, mean, na.rm=TRUE)
```
##Plotting the time-series graph

```{r}
plot(Intervals, StepsPerIntervalaverage, type="l",
     main="Average number of steps per interval across all days", 
     xlab="Interval", ylab="Average # of steps across all days", 
     lwd=2, col="yellow")
```
##The location of where the maximum is

```{r}
maxIntervalDays <- max(StepsPerIntervalaverage, na.rm=TRUE)
maxIndex <- as.numeric(which(StepsPerIntervalaverage == maxIntervalDays))
maxIndex
```
#Plotting a vertical line where the max is

```{r}
maxInterval <- Intervals[maxIndex]
abline(v=maxInterval, col="red", lwd=3)

maxInterval
```
## Part4- Calculating total amount of missing values in the data set


```{r}
completeRows <- complete.cases(activity$steps)
numNA <- sum(as.numeric(!completeRows))
numNA

StepsPerDaymean[is.nan(StepsPerDaymean)] <- 0
meanColumn <- rep(StepsPerDaymean, 288)
rawSteps <- activity$steps
stepsNA <- is.na(rawSteps)
rawSteps[stepsNA] <- meanColumn[stepsNA]
datNew <- activity
datNew$steps <- rawSteps

stepsNew <- split(datNew$steps, dates$yday)
```
##Finding the total number of steps over each day

```{r}
totalStepsPerDayNew <- sapply(stepsNew, sum)
```
#Plotting histogram 

```{r}
par(mfcol=c(2,1))
```
#Plot the original histogram first

```{r}
plot(Dates, totalStepsPerDay, main="Histogram of steps taken each day before imputing", 
     xlab="Date (October to November 2012)", ylab="Frequency", type="h", lwd=4, col="pink")
```
#Plot the modified histogram after

```{r}
plot(Dates, totalStepsPerDayNew, main="Histogram of steps taken each day after imputing", 
     xlab="Date (October to November 2012)", ylab="Frequency", type="h", lwd=4, col="pink")
```
##New mean

```{r}
StepsPerDaymeanNew <- sapply(stepsNew, mean)
DataFramemeanNew <- data.frame(date=Dates, StepsPerDaymean=StepsPerDaymean, 
                               StepsPerDaymeanNew=StepsPerDaymeanNew, row.names=NULL)
DataFramemeanNew

```
##New median

```{r}
StepsPerDaymedianNew <- sapply(stepsNew, median)
DataFramemedianNew <- data.frame(date=Dates, StepsPerDaymedian=StepsPerDaymedian, 
                                 StepsPerDaymedianNew=StepsPerDaymedianNew, row.names=NULL)
DataFramemedianNew
```


## Part5- Splitting up the data so that it's sorted by weekday or weekend
##0 is for Sunday, 1 is for Monday, going up to 6 for Saturday

```{r}
wdays <- dates$wday
classifywday <- rep(0, 17568) # 17568 observations overall
classifywday[wdays >= 1 & wdays <= 5] <- 1
classifywday[wdays == 6 | wdays == 0] <- 2

```
#Create a new factor variable that has labels Weekdays and Weekends
```{r}
daysFactor <- factor(classifywday, levels=c(1,2), labels=c("Weekdays", "Weekends"))
```
#Create a new column that contains this factor for each day

```{r}
datNew$typeOfDay <- daysFactor
```

#Splitting up into two data frames

```{r}
Weekdays <- datNew[datNew$typeOfDay == "Weekdays", ]
Weekends <- datNew[datNew$typeOfDay == "Weekends", ]


SplitWeekdays <- split(Weekdays$steps, Weekdays$interval)
SplitWeekends <- split(Weekends$steps, Weekends$interval)

```
#Finding the average for each interval

```{r}
StepsPerWeekdaymeanInterval <- sapply(SplitWeekdays, mean)
StepsPerWeekendmeanInterval <- sapply(SplitWeekends, mean)

par(mfcol=c(2,1))
plot(Intervals, StepsPerWeekdaymeanInterval, type="l",
     main="Average number of steps per interval across all weekdays", 
     xlab="Interval", ylab="Average # of steps across all weekdays", 
     lwd=2, col="green")

plot(Intervals, StepsPerWeekendmeanInterval, type="l",
     main="Average number of steps per interval across all weekends", 
     xlab="Interval", ylab="Average # of steps across all weekends", 
     lwd=2, col="green")
```



Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
